Raw Lending Club Analysis
========================

------

### Load libraries

```{r}
library(Hmisc)
```


------

## Processing

Download the data, read the data in and save the raw data as an RDA file

```{r loadData,cache=TRUE}
getwd()
download.file("https://spark-public.s3.amazonaws.com/dataanalysis/loansData.rda",destfile="../../data/loansData.rda")
dateDownloaded <- date()
dateDownloaded
load('../../data/loansData.rda')
```


### Look at the data set

```{r, dependson="loadData"}
dim(loansData)
head(loansData)
summary(loansData)
sapply(loansData[1,],class)
```

*Interest.Rate* and *Debt.To.Income.Ratio* variables have to be transformed from `factor` type to `numeric type`


### Fix columns that have wrong variable types
```{r, dependson="loadData"}
loansData <- transform(loansData, Interest.Rate = as.numeric(sub("%","",Interest.Rate)),
                       Debt.To.Income.Ratio = as.numeric(sub("%","",Debt.To.Income.Ratio)))
```


### Find out about missing values and take out imcomplete cases

```{r, dependson="loadData"}
table(is.na(loansData))
loansData <- loansData[complete.cases(loansData),]
```


### Add a variable corresponding the percentage of the requested amount that was finally loaned to the applicant

```{r, dependson="loadData"}
Percent.Funded <- round(100*(1 - (loansData$Amount.Requested-loansData$Amount.Funded.By.Investors)/loansData$Amount.Requested), 2)
loansData$Percent.Funded <- Percent.Funded
```



------

## Exploratory analysis

### Make some univariate tables

```{r, dependson="loadData"}
length(unique(loansData$Interest.Rate))
length(unique(loansData$Debt.To.Income.Ratio))
length(unique(loansData$Loan.Purpose))
length(unique(loansData$FICO.Range))
length(unique(loansData$Employment.Length))
table(loansData$Loan.Purpose)
table(loansData$State)
table(loansData$FICO.Range)
table(loansData$Employment.Length)
```


### Make some univariate plots/summaries

```{r, dependson="loadData"}
hist(loansData$Amount.Requested,col="blue")
quantile(loansData$Amount.Requested)
```

```{r, dependson="loadData"}
hist(loansData$Amount.Funded.By.Investors,col="blue")
quantile(loansData$Amount.Funded.By.Investors)
```

```{r, dependson="loadData"}
hist(loansData$Interest.Rate,col="blue")
quantile(loansData$Interest.Rate)
```

```{r, dependson="loadData"}
hist(loansData$Debt.To.Income.Ratio,col="blue")
quantile(loansData$Debt.To.Income.Ratio)
```

```{r, dependson="loadData"}
hist(loansData$Monthly.Income,col="blue")
quantile(loansData$Monthly.Income)
plot(loansData$Monthly.Income,col="blue")
hist(loansData$Monthly.Income[loansData$Monthly.Income < 39000],col="blue",breaks=90)
loansData <- loansData[loansData$Monthly.Income < 39000,]
```

3 outliers among *Monthly.Income* values

```{r, dependson="loadData"}
hist(loansData$Open.CREDIT.Lines,col="blue",breaks=50)
quantile(loansData$Open.CREDIT.Lines)
```

```{r, dependson="loadData"}
hist(loansData$Revolving.CREDIT.Balance,col="blue")
quantile(loansData$Revolving.CREDIT.Balance)
plot(loansData$Revolving.CREDIT.Balance,col="blue")
hist(loansData$Revolving.CREDIT.Balance[loansData$Revolving.CREDIT.Balance < 150000],col="blue",breaks=60)
loansData <- loansData[loansData$Revolving.CREDIT.Balance < 150000,]
```

6 outliers among *Revolving.Credit.Balance* values

```{r, dependson="loadData"}
hist(loansData$Inquiries.in.the.Last.6.Months,col="blue")
quantile(loansData$Inquiries.in.the.Last.6.Months)
```

```{r, dependson="loadData"}
hist(loansData$Percent.Funded,col="blue",breaks=100)
quantile(loansData$Percent.Funded)
```


### Calculate percentage of loans that were 100% funded (or almost 100%) and those which were not funded at all

```{r, dependson="loadData"}
100*sum(loansData$Percent.Funded > 98)/nrow(loansData)
100*sum(loansData$Percent.Funded == 0)/nrow(loansData)
```


### Plot interest rate versus FICO range

```{r, dependson="loadData"}
plot(loansData$FICO.Range,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

A clear association between *FICO.Range* and *Interest.Rate*. No big surprise here.

### Plot interest rate versus debt to income ratio
```{r, dependson="loadData"}
plot(loansData$Debt.To.Income.Ratio,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

No clear association between *Debt.To.Income.Ratio* and *Interest.Rate*.

### Plot interest rate versus amount requested
```{r, dependson="loadData"}
plot(loansData$Amount.Requested,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
lm0 <- lm(loansData$Interest.Rate ~ loansData$Amount.Requested)
lines(loansData$Amount.Requested,lm0$fitted,col="red",lwd=3)
```

A not so strong association between *Amount.Requested* and *Interest.Rate*.

### Plot interest rate versus loan length
```{r, dependson="loadData"}
plot(loansData$Loan.Length,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

A clear association between *Loan.Length* and *Interest.Rate*.

### Plot interest rate versus monthly income
```{r, dependson="loadData"}
plot(loansData$Monthly.Income,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

No clear association between *Monthly.Income* and *Interest.Rate*.

### Plot interest rate versus open credit lines
```{r, dependson="loadData"}
plot(loansData$Open.CREDIT.Lines,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

No clear association between *Open.CREDIT.Lines* and *Interest.Rate*.

### Plot interest rate versus revolving credit balance
```{r, dependson="loadData"}
plot(loansData$Revolving.CREDIT.Balance,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

No clear association between *Revolving.CREDIT.Balance* and *Interest.Rate*.

### Plot interest rate versus employment length
```{r, dependson="loadData"}
plot(loansData$Employment.Length,loansData$Interest.Rate,pch=19,col="blue",cex=0.5)
```

No clear association between *Employment.Length* and *Interest.Rate*.


------

## Modeling


## Fit a basic model relating interest rate to FICO score

```{r, dependson="loadData"}
lm1 <- lm(loansData$Interest.Rate ~ loansData$FICO.Range)
summary(lm1)
```


## Plot the observed ('black') and fitted ('red') points

```{r, dependson="loadData"}
plot(loansData$FICO.Range,loansData$Interest.Rate,pch=19)
points(loansData$FICO.Range,lm1$fitted,pch=19,col="red")
```


## Look at residuals versus observations, residuals versus fitted values

```{r, dependson="loadData"}
par(mfrow=c(1,2))
plot(loansData$FICO.Range,lm1$residuals,pch=19)
plot(lm1$fitted,lm1$residuals,pch=19)
```



## Fit a basic model relating interest rate to FICO score and amount requested

```{r, dependson="loadData"}
lm2 <- lm(loansData$Interest.Rate ~ loansData$Amount.Requested)
summary(lm2)
plot(lm2$fitted,lm2$residuals,pch=19)
```


## Fit a basic model relating interest rate to FICO score and amount requested

```{r, dependson="loadData"}
lm3 <- lm(loansData$Interest.Rate ~ loansData$FICO.Range + loansData$Amount.Requested)
summary(lm3)
plot(lm3$fitted,lm3$residuals,pch=19)
```


## Fit a basic model relating interest rate to FICO score and loan length

```{r, dependson="loadData"}
lm4 <- lm(loansData$Interest.Rate ~ loansData$FICO.Range + loansData$Loan.Length)
summary(lm4)
```


## Fit a basic model relating interest rate to FICO score, amount requested and loan length

```{r, dependson="loadData"}
lm5 <- lm(loansData$Interest.Rate ~ loansData$Amount.Requested + loansData$Loan.Length + loansData$FICO.Range)
summary(lm5)
```

## Fit a basic model relating interest rate to FICO score, amount requested and loan length

```{r, dependson="loadData"}
lm10 <- lm(loansData$Interest.Rate ~ loansData$Loan.Purpose)
summary(lm10)
```

## Let's use model 4

```{r, dependson="lm5Chunk"}
summary(lm4)
confint(lm4)


par(mfrow=c(2,2))

hist(loansData$Amount.Requested,col="blue",breaks=20,main="",xlab="Amount requested (in dollars)")
plot(loansData$FICO.Range,loansData$Interest.Rate,pch=19,col="blue",cex=0.5,xlab="FICO score range",ylab="Interest rate (%)")
plot(lm2$fitted,lm2$residuals,pch=19,col=loansData$Loan.Length, xlab="Interest rate (%)", ylab="No Adjustment Residuals")
plot(lm4$fitted,lm4$residuals,pch=19,col=loansData$Loan.Length, xlab="Interest rate (%)", ylab="Adjusted Residuals")


par(mfrow=c(2,3))

plot(loansData$Amount.Requested,loansData$Interest.Rate,pch=19,col="blue",cex=0.5,xlab="Amount Requested (in dollars)",ylab="Interest rate (%)")
lm0 <- lm(loansData$Interest.Rate ~ loansData$Amount.Requested)
lines(loansData$Amount.Requested,lm0$fitted,col="red",lwd=3)

plot(loansData$FICO.Range,loansData$Interest.Rate,pch=19,col="blue",cex=0.5,xlab="FICO score range",ylab="Interest rate (%)")

plot(loansData$Loan.Length,loansData$Interest.Rate,pch=19,col="blue",cex=0.5,xlab="Loan Length (in months)",ylab="Interest rate (%)")

plot(lm2$fitted,lm2$residuals,pch=19,col=loansData$Loan.Length, xlab="Interest rate (%)", ylab="No Adjustment Residuals")
plot(lm5$fitted,lm5$residuals,pch=19,col=loansData$Loan.Length, xlab="Interest rate (%)", ylab="Adjusted Residuals")


```



