Samsung Activity Prediction Study
=================================

---
title: "Samsung Activity Prediction Study"
output:
  html_document:
    keep_md: true
---

------

### Load libraries

```{r}
library(tree)
library(randomForest)
```


------

## Processing

Download the data, read the data in and save the raw data as an RDA file

```{r loadData,cache=TRUE}
#getwd()
#download.file("https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda",destfi#le="../../data/samsungData.rda")
#dateDownloaded <- date()
#dateDownloaded
load('../data/samsungData.rda')
```


### Look at the data set

```{r, dependson="loadData"}
dim(samsungData)
samsungData <- data.frame(samsungData)
names(samsungData)[duplicated(names(samsungData))] # 84 duplicate column names found!
samsungData <- data.frame(samsungData) # This removes duplicates
names(samsungData) <- c(paste0("V",c(1:561)),"subject","activity") # this should also solve it!
table(samsungData$activity)
sapply(samsungData[1,],class)
samsungData$activity <- sapply(samsungData$activity, factor) # set 'activity' variable to be a factor
table(samsungData$subject)
length(unique(samsungData$subject))
```


### Find out about missing values and take out imcomplete cases

```{r, dependson="loadData"}
table(is.na(samsungData))
```

### Building training and test sets and make necessary transformations

```{r, dependson="loadData"}
trainData <- samsungData[samsungData$subject %in% c(1,3,5,6,7,8,11,14,15,16,17,19,21,22,23,25,26),]
trainData <- trainData[,-562] # Eliminate subject variable
testData <- samsungData[samsungData$subject %in% c(27,28,29,30),]
testData <- testData[,-562] # Eliminate subject variable
```


------

## Exploratory Analysis

### Plotting average acceleration for first subject

```{r, dependson="loadData"}
par(mfrow=c(1,2))
numericActivity <- as.numeric(as.factor(samsungData$activity))[samsungData$subject==1]
plot(samsungData[samsungData$subject==1,1], pch=19, col=numericActivity, ylab=names(samsungData)[1])
plot(samsungData[samsungData$subject==1,2], pch=19, col=numericActivity, ylab=names(samsungData)[2])
legend(150, -0.1, legend=unique(samsungData$activity), col=unique(numericActivity), pch=19)
```


### Plotting max acceleration for first subject

```{r, dependson="loadData"}
par(mfrow=c(1,2))
plot(samsungData[samsungData$subject==1,10], pch=19, col=numericActivity, ylab=names(samsungData)[10])
plot(samsungData[samsungData$subject==1,11], pch=19, col=numericActivity, ylab=names(samsungData)[11])
```



------

## Predictive Modeling


## Build a predictive tree model from the training data set

```{r, dependson="loadData"}
tree1 <- tree(activity ~ ., data=trainData)
summary(tree1)
plot(tree1)
text(tree1)
```


## Cross-validate tree model

```{r, dependson="loadData"}
plot(cv.tree(tree1,FUN=prune.tree,method="misclass"))
plot(cv.tree(tree1))
```


## Prune the tree

```{r, dependson="loadData"}
pruneTree <- prune.tree(tree1, best=6)
summary(pruneTree)
plot(pruneTree)
text(pruneTree)
```


## Show resubstitution error

```{r, dependson="loadData"}
table(trainData$activity, predict(pruneTree,type="class"))
table(trainData$activity, predict(tree1,type="class"))
sum(predict(pruneTree,type="class") != trainData$activity)/length(trainData$activity)
sum(predict(tree1,type="class") != trainData$activity)/length(trainData$activity)
```


## Random forest

```{r, dependson="loadData"}
forest1 <- randomForest(activity ~ ., data=trainData, prox=TRUE)
forest1
sum(predict(forest1, type="class") != trainData$activity)/length(trainData$activity)
```


## Prediction with test set

```{r, dependson="loadData"}
sum(predict(tree1, testData, type="class") != testData$activity)/length(testData$activity)
sum(predict(pruneTree, testData, type="class") != testData$activity)/length(testData$activity)
sum(predict(forest1, testData, type="class") != testData$activity)/length(testData$activity)
```

## Final graphs

```{r, dependson="loadData"}
par(mfrow=c(2,2))
plot(tree1)
text(tree1)
plot(pruneTree)
text(pruneTree)
plot(cv.tree(tree1,FUN=prune.tree,method="misclass"))
plot(cv.tree(tree1))
```

